---
title: "Homework 2"
author: Ben Lebwohl
output: github_document
---

Load libraries
```{r}
library(tidyverse)
library(readxl)
```

## Problem 1

```{r}
trash_wheel_df = 
read_excel("./data_files/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range= "A2:N406") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) 
trash_wheel_df
```

> Round the number of sports balls to the nearest integer and converts the result to an integer variable 

```{r}
trash_wheel_df =
trash_wheel_df %>% 
  mutate (
  sports_balls = as.integer(sports_balls)
  )
trash_wheel_df
```


> Read and clean precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. 

```{r}
precip_2017 = 
read_excel("./data_files/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", range= "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(total >0) %>% 
  mutate(year = "2017")

precip_2018 = 
read_excel("./data_files/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range= "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(total >0) %>% 
  mutate(year = "2018")
```

> Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
combined_precip = 
bind_rows(precip_2017, precip_2018) %>%
  mutate(month = month.name[month]) %>%
  relocate(month, year)
combined_precip
```

> For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

The total precipitation in 2018 was `r sum(precip_2018$total)` inches.


The median number of sports balls in a dumpster in 2017 was  `r median(trash_wheel_df[year=2017]$sports_balls)`.

## Problem 2

>Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r}
transit_df = 
read_csv("./data_files/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>% 
  mutate(entry = ifelse(entry== "YES",T,F))
```

> Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?


This data set contains a list of subway stations and their associated characteristics, including location (longitude and latitude), whether entry is available, and whether it is ADA-compliant. It is an 1868 by 19 dataset. It is not tidy, because the routes are listed as separate columns. 

_Answer the following questions using these data:_

> How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.

```{r}
nrow(distinct(transit_df,station_name))
```


> How many stations are ADA compliant?

Create a data set consisting of only distinct stations, and then count the number that ADA compliant:

```{r}
distinct_stations_df = 
  transit_df %>%
  distinct(station_name, .keep_all=TRUE)
table(distinct_stations_df$ada)
```

> What proportion of station entrances / exits without vending allow entrance?

Create a data set consisting of stations without vending, and then calculate the proportion that allow entrance:
```{r}
non_vending_df=
transit_df %>%
  filter(vending == "NO")
prop.table(table(non_vending_df$entry))
```

>Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
transit_df_tidy = 
  transit_df %>%
  mutate_at(
    c("route8","route9","route10","route11"), as.character
    ) %>% 
  pivot_longer(
    cols = route1:route11, 
    names_prefix = "route",
    values_to = "route",
    values_drop_na = TRUE
  )
transit_df_tidy =
  select(transit_df_tidy,-name)
  
```

The number of distinct stations that serve the A train:
```{r}
distinct_transit_df_tidy =
  transit_df_tidy %>%
  distinct(station_name, .keep_all=TRUE)
table(distinct_transit_df_tidy$route)
```

48 stations serve the A train


The number of A train stations that are ADA compliant:

```{r}
a_train_stations_df =
  distinct_transit_df_tidy %>% 
  filter(route =="A")
table(a_train_stations_df$ada)
```

## Problem 3

>First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols_df=
read_csv("./data_files/fivethirtyeight_datasets/pols-month.csv") %>% 
separate("mon", c("year", "month", "day"), sep = "-") %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate (pres= ifelse(prez_gop==1,"gop","dem"))
pols_df=  
select(pols_df, -prez_dem, -prez_gop, -day)
```

>Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp_df =
  read_csv("./data_files/fivethirtyeight_datasets/snp.csv") %>% 
separate("date", c("month", "day", "year"), sep = "/") %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month])
snp_df = select(snp_df, -day)
snp_df= select(snp_df, year, everything())
snp_df= arrange(snp_df,year)
```

>Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.


```{r}
unemployment_df =
  read_csv("./data_files/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec, 
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(month = match(month, month.abb)) %>% 
  mutate(month = month.name[month])
```

>Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
merged538_df =
  left_join(pols_df, snp_df, by =c("year","month"))
```

Convert the year variable in unemployment_df to a character variable
```{r}
  unemployment_df =
  mutate(unemployment_df, year = as.character(year)
  )
```

Merge the unemoloyment data into the main data set
```{r}
merged538_df =
  left_join(unemployment_df, merged538_df, by = c("year","month"))
merged538_df= select(merged538_df, year, month, unemployment, close, pres, everything())
merged538_df
```

  

